---
title: "TASK1"
author: "Dilrukshi KAwya"
output:
  pdf_document: default
  html_document: default
---

```{r}
setwd("C:/codesoft/TASK1_CODESOFT")
getwd()
```
```{r}

titanic<-read.csv("Titanic-Dataset.csv")
str(titanic)
summary(titanic)
```

#check the missing values

```{r}
sapply(titanic,function(x) sum(is.na(x)))
```
#fill missing values in the Age column with the mean Age   



```{r}
titanic$Age[is.na(titanic$Age)] <- mean(titanic$Age,na.rm = TRUE)

titanic$Embarked[is.na(titanic$Embarked)]<-"S"
```

#convert categorical variables to factors   

```{r}
titanic$Sex<-as.factor(titanic$Sex)
titanic$Embarked<-as.factor(titanic$Embarked)

```

#drop un necessary columns    

```{r}
titanic<-titanic%>%select(-Name,-Ticket,-Cabin)
head(titanic)
```

```{r}
titanic$Familysize<- titanic$SibSp + titanic$Parch + 1
test_data$Familysize <- test_data$SibSp + test_data$Parch + 1

```

#split the data partition   

```{r}
set.seed(123)
titanicIndex<-createDataPartition(titanic$Survived,p=0.8,list = FALSE)
train_data<- titanic[titanicIndex,]
test_data<- titanic[-titanicIndex,]
```

#Build a RandomForest Model   

```{r}
model<-randomForest(Survived ~ Pclass + Sex + Age + Fare + Embarked + Familysize,data = train_data)
print(model)
```

#predicting on Test Data   
```{r}

test_pred<-predict(model,newdata = test_data)
test_pred
```
# Evaluate the Model   

```{r}
test_data$Survived <-factor(test_data$Survived, levels = c(0, 1))
test_pred <- factor(test_pred, levels = levels(test_data$Survived))
levels(test_pred) <- levels(test_data$Survived)


confusionMatrix(test_pred,test_data$Survived)
```

```{r}
importance<- varImpPlot(model)
print(importance)
```
```{r}
new_data <- data.frame(
  Pclass = 3,
  Sex = factor("male", levels = levels(titanic$Sex)),
  Age = 22,
  SibSp = 1,
  Parch = 0,
  Fare = 7.25,
  Familysize = 2,
  Embarked = factor("S", levels = levels(titanic$Embarked))
)

new_pred<- predict(model,new_data)
print(new_pred)


```
#Summary    
 Data Preparation: The Titanic dataset (titanic) is split into training and test sets.   
 
Model Training: A random forest classifier (randomForest) is trained on the training data (train_data).   
Model Evaluation: The model's performance is evaluated using a confusion matrix (confusionMatrix) on the test data (test_data).    

Feature Importance: The importance of each predictor variable is visualized (varImpPlot).    

Prediction: The model is used to predict survival for a new passenger (new_data).    

#interpretation    





